---
title: "2.0 TAI profiles"
output:
  html_document:
    fig_width: 6
    fig_height: 4
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

# TAI profiles

From PES, one can obtain the TAI profiles through the `mytTAI::TAI` function. The `myTAI` package, to my knowledge, does not enable the flexibility to plot the separate sexes, for example in one plot. Therefore, I will _not_ use the `myTAI::PlotSignature` function.

Here, I show the TAI profiles for the developmental stages. For tissues of the adult (matSP) _Fucus_ species, one can find them in `2.1 TAI tissues in Fucus matSP`.

```{r}
library(tidyverse)
library(myTAI)
```

## Load PES

Non-transformed

```{r}
Fd_PES <-
  readr::read_csv(file = "data/Fd_PES.csv")
Fd_PES_F <-
  readr::read_csv(file = "data/Fd_PES_F.csv")
Fd_PES_M <-
  readr::read_csv(file = "data/Fd_PES_M.csv")
Fs_PES <-
  readr::read_csv(file = "data/Fs_PES.csv")
Fs_PES_F <-
  readr::read_csv(file = "data/Fs_PES_F.csv")
Fs_PES_M <-
  readr::read_csv(file = "data/Fs_PES_M.csv")
Ec_PES_32m <-
  readr::read_csv(file = "data/Ec_PES_32m.csv")
Ec_PES_25f <-
  readr::read_csv(file = "data/Ec_PES_25f.csv")
```

sqrt-tranformed

```{r}
Fd_PES.sqrt <-
  readr::read_csv(file = "data/Fd_PES.sqrt.csv")
Fd_PES_F.sqrt <-
  readr::read_csv(file = "data/Fd_PES_F.sqrt.csv")
Fd_PES_M.sqrt <-
  readr::read_csv(file = "data/Fd_PES_M.sqrt.csv")
Fs_PES.sqrt <-
  readr::read_csv(file = "data/Fs_PES.sqrt.csv")
Fs_PES_F.sqrt <-
  readr::read_csv(file = "data/Fs_PES_F.sqrt.csv")
Fs_PES_M.sqrt <-
  readr::read_csv(file = "data/Fs_PES_M.sqrt.csv")
Ec_PES_32m.sqrt <-
  readr::read_csv(file = "data/Ec_PES_32m.sqrt.csv")
Ec_PES_25f.sqrt <-
  readr::read_csv(file = "data/Ec_PES_25f.sqrt.csv")
```

log2-tranformed

```{r}
Fd_PES.log2 <-
  readr::read_csv(file = "data/Fd_PES.log2.csv")
Fd_PES_F.log2 <-
  readr::read_csv(file = "data/Fd_PES_F.log2.csv")
Fd_PES_M.log2 <-
  readr::read_csv(file = "data/Fd_PES_M.log2.csv")
Fs_PES.log2 <-
  readr::read_csv(file = "data/Fs_PES.log2.csv")
Fs_PES_F.log2 <-
  readr::read_csv(file = "data/Fs_PES_F.log2.csv")
Fs_PES_M.log2 <-
  readr::read_csv(file = "data/Fs_PES_M.log2.csv")
Ec_PES_32m.log2 <-
  readr::read_csv(file = "data/Ec_PES_32m.log2.csv")
Ec_PES_25f.log2 <-
  readr::read_csv(file = "data/Ec_PES_25f.log2.csv")
```

rank-tranformed

```{r}
Fd_PES.rank <-
  readr::read_csv(file = "data/Fd_PES.rank.csv")
Fd_PES_F.rank <-
  readr::read_csv(file = "data/Fd_PES_F.rank.csv")
Fd_PES_M.rank <-
  readr::read_csv(file = "data/Fd_PES_M.rank.csv")
Fs_PES.rank <-
  readr::read_csv(file = "data/Fs_PES.rank.csv")
Fs_PES_F.rank <-
  readr::read_csv(file = "data/Fs_PES_F.rank.csv")
Fs_PES_M.rank <-
  readr::read_csv(file = "data/Fs_PES_M.rank.csv")
Ec_PES_32m.rank <-
  readr::read_csv(file = "data/Ec_PES_32m.rank.csv")
Ec_PES_25f.rank <-
  readr::read_csv(file = "data/Ec_PES_25f.rank.csv")
```

rlog-tranformed

```{r}
Fd_PES.rlog <-
  readr::read_csv(file = "data/Fd_PES.rlog.csv")
Fd_PES_F.rlog <-
  readr::read_csv(file = "data/Fd_PES_F.rlog.csv")
Fd_PES_M.rlog <-
  readr::read_csv(file = "data/Fd_PES_M.rlog.csv")
Fs_PES.rlog <-
  readr::read_csv(file = "data/Fs_PES.rlog.csv")
Fs_PES_F.rlog <-
  readr::read_csv(file = "data/Fs_PES_F.rlog.csv")
Fs_PES_M.rlog <-
  readr::read_csv(file = "data/Fs_PES_M.rlog.csv")
Ec_PES_32m.rlog <-
  readr::read_csv(file = "data/Ec_PES_32m.rlog.csv")
Ec_PES_25f.rlog <-
  readr::read_csv(file = "data/Ec_PES_25f.rlog.csv")
```

## myTAI issue

I will post an issue on GitHub but `myTAI::tf()` loses the column name hwen transforming only one column.

# TAI calculation

I had previously used a function to combine males and females into one TAI plot.

```{r}
#TI stands for transcriptome index
TI.preplot <- function(ExpressionSet, permutations = 50000){
  std <- 
    myTAI::bootMatrix(
      ExpressionSet = ExpressionSet, 
      permutations  = permutations) %>% 
    apply(2, stats::sd)
  TI.out <- 
    myTAI::TAI(ExpressionSet) %>%
    tibble::as_tibble(rownames = "Stage", colnames = c("TAI", "PS")) %>% 
    dplyr::bind_cols(as_tibble(std)) %>%
    dplyr::rename(TAI = 2, sd = 3)
  return(TI.out)
}
```

### Ectocarpus

```{r}
# function to process Fd. This will be changed to accommodate more seq data.
get_TAI_Ec <- function(PES_M, PES_F, ordered_stages){
  TAI_M <- 
    TI.preplot(PES_M) %>%
    dplyr::mutate(Sex = "Male")
  TAI_F <- 
    TI.preplot(PES_F) %>%
    dplyr::mutate(Sex = "Female")
  stages_to_NA <- # this will differ for Fucus distichus
    ordered_stages[-c(1, 1+1, length(ordered_stages) - 1, length(ordered_stages))]
  TAI_out <- 
    dplyr::bind_rows(TAI_M, TAI_F)
  TAI_out$Stage <- base::factor(TAI_out$Stage, ordered_stages)
  return(TAI_out)
}
```

```{r}
ordered_stages <- colnames(Ec_PES_25f)[3:ncol(Ec_PES_25f)]

Ec_TAI <- 
  get_TAI_Ec(
    PES_M = Ec_PES_32m, 
    PES_F = Ec_PES_25f, 
    ordered_stages)

Ec_TAI.sqrt <- 
  get_TAI_Ec(
    PES_M = Ec_PES_32m.sqrt, 
    PES_F = Ec_PES_25f.sqrt, 
    ordered_stages)

Ec_TAI.log2 <- 
  get_TAI_Ec(
    PES_M = Ec_PES_32m.log2, 
    PES_F = Ec_PES_25f.log2, 
    ordered_stages)

Ec_TAI.rank <- 
  get_TAI_Ec(
    PES_M = Ec_PES_32m.rank, 
    PES_F = Ec_PES_25f.rank, 
    ordered_stages)

Ec_TAI.rlog <- 
  get_TAI_Ec(
    PES_M = Ec_PES_32m.rlog, 
    PES_F = Ec_PES_25f.rlog, 
    ordered_stages)
```


# TAI profiles

## Ectocarpus

```{r}
Ec_TAI %>% 
  ggplot2::ggplot(
    aes(
      y = TAI, 
      x = Stage,
      group = Sex,
      colour = Sex,
      fill = Sex)) +
  ggplot2::geom_line(alpha = 0.4, linewidth = 1) +
  ggplot2::geom_crossbar(
    aes(ymin = TAI - sd,
        ymax = TAI + sd
        ), 
    alpha = 0.5,
    position = ggplot2::position_dodge(0.55), 
    width = 0.5) +
  theme_classic() +
  ggsci::scale_colour_npg() +
  ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge=2)) +
  labs(title = "Ectocarpus",
       subtitle = "TPM")

Ec_TAI.sqrt %>% 
  ggplot2::ggplot(
    aes(
      y = TAI, 
      x = Stage,
      group = Sex,
      colour = Sex,
      fill = Sex)) +
  ggplot2::geom_line(alpha = 0.4, linewidth = 1) +
  ggplot2::geom_crossbar(
    aes(ymin = TAI - sd,
        ymax = TAI + sd
        ), 
    alpha = 0.5,
    position = ggplot2::position_dodge(0.55), 
    width = 0.5) +
  theme_classic() +
  ggsci::scale_colour_npg() +
  ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge=2)) +
  labs(title = "Ectocarpus",
       subtitle = "sqrt(TPM)")

Ec_TAI.log2 %>% 
  ggplot2::ggplot(
    aes(
      y = TAI, 
      x = Stage,
      group = Sex,
      colour = Sex,
      fill = Sex)) +
  ggplot2::geom_line(alpha = 0.4, linewidth = 1) +
  ggplot2::geom_crossbar(
    aes(ymin = TAI - sd,
        ymax = TAI + sd
        ), 
    alpha = 0.5,
    position = ggplot2::position_dodge(0.55), 
    width = 0.5) +
  theme_classic() +
  ggsci::scale_colour_npg() +
  ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge=2)) +
  labs(title = "Ectocarpus",
       subtitle = "log2(TPM+\u03b1)")

Ec_TAI.rank %>% 
  ggplot2::ggplot(
    aes(
      y = TAI, 
      x = Stage,
      group = Sex,
      colour = Sex,
      fill = Sex)) +
  ggplot2::geom_line(alpha = 0.4, linewidth = 1) +
  ggplot2::geom_crossbar(
    aes(ymin = TAI - sd,
        ymax = TAI + sd
        ), 
    alpha = 0.5,
    position = ggplot2::position_dodge(0.55), 
    width = 0.5) +
  theme_classic() +
  ggsci::scale_colour_npg() +
  ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge=2)) +
  labs(title = "Ectocarpus",
       subtitle = "rank(TPM)")

Ec_TAI.rlog %>% 
  ggplot2::ggplot(
    aes(
      y = TAI, 
      x = Stage,
      group = Sex,
      colour = Sex,
      fill = Sex)) +
  ggplot2::geom_line(alpha = 0.4, linewidth = 1) +
  ggplot2::geom_crossbar(
    aes(ymin = TAI - sd,
        ymax = TAI + sd
        ), 
    alpha = 0.5,
    position = ggplot2::position_dodge(0.55), 
    width = 0.5) +
  theme_classic() +
  ggsci::scale_colour_npg() +
  ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge=2)) +
  labs(title = "Ectocarpus",
       subtitle = "rlog(TPM)")
```

The r-log is behaving weird because the transformation is non-linear but also does not preserve the _monotonic_ relationship.

Get session info.

```{r}
devtools::session_info()
```