---
title: "2.1 TAI tissues in Fucus matSP"
output:
  html_document:
    fig_width: 6
    fig_height: 4
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

# TAI profiles of tissues

For TAI profiles for the developmental stages, one can find them in `2.0 TAI profiles`.

```{r}
library(tidyverse)
library(myTAI)
```

Non-transformed

```{r}
# tissues in the matSP
Fd_PES_matSP <-
  readr::read_csv(file = "data/Fd_PES_matSP.csv")
Fs_PES_F_matSP <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.csv")
Fs_PES_M_matSP <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.csv")
```

sqrt-tranformed

```{r}
# tissues in the matSP
Fd_PES_matSP.sqrt <-
  readr::read_csv(file = "data/Fd_PES_matSP.sqrt.csv")
Fs_PES_F_matSP.sqrt <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.sqrt.csv")
Fs_PES_M_matSP.sqrt <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.sqrt.csv")
```

log2-tranformed

```{r}
# tissues in the matSP
Fd_PES_matSP.log2 <-
  readr::read_csv(file = "data/Fd_PES_matSP.log2.csv")
Fs_PES_F_matSP.log2 <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.log2.csv")
Fs_PES_M_matSP.log2 <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.log2.csv")
```

rank-tranformed

```{r}
# tissues in the matSP
Fd_PES_matSP.rank <-
  readr::read_csv(file = "data/Fd_PES_matSP.rank.csv")
Fs_PES_F_matSP.rank <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.rank.csv")
Fs_PES_M_matSP.rank <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.rank.csv")
```

rlog-tranformed

```{r}
# tissues in the matSP
Fd_PES_matSP.rlog <-
  readr::read_csv(file = "data/Fd_PES_matSP.rlog.csv")
Fs_PES_M_matSP.rlog <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.rlog.csv")
Fs_PES_F_matSP.rlog <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.rlog.csv")
```

# Tissue-specific TAI

```{r}
#TI stands for transcriptome index
TI.preplot <- function(ExpressionSet, permutations = 2000){
  std <- 
    myTAI::bootMatrix(
      ExpressionSet = ExpressionSet, 
      permutations  = permutations) %>% 
    apply(2, stats::sd)
  TI.out <- 
    myTAI::TAI(ExpressionSet) %>%
    tibble::as_tibble(rownames = "Tissue", colnames = c("TAI", "PS")) %>% 
    dplyr::bind_cols(as_tibble(std)) %>%
    dplyr::rename(TAI = 2, sd = 3)
  return(TI.out)
}
```

```{r}
Fd_TAI_matSP <- 
  TI.preplot(Fd_PES_matSP)
Fs_TAI_F_matSP <- 
  TI.preplot(Fs_PES_F_matSP) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP <- 
  TI.preplot(Fs_PES_M_matSP) %>%
  dplyr::mutate(Sex = "M")

# sqrt
Fd_TAI_matSP.sqrt <- 
  TI.preplot(Fd_PES_matSP.sqrt)
Fs_TAI_F_matSP.sqrt <- 
  TI.preplot(Fs_PES_F_matSP.sqrt) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.sqrt <- 
  TI.preplot(Fs_PES_M_matSP.sqrt) %>%
  dplyr::mutate(Sex = "M")

# log2
Fd_TAI_matSP.log2 <- 
  TI.preplot(Fd_PES_matSP.log2)
Fs_TAI_F_matSP.log2 <- 
  TI.preplot(Fs_PES_F_matSP.log2) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.log2 <- 
  TI.preplot(Fs_PES_M_matSP.log2) %>%
  dplyr::mutate(Sex = "M")

# rank
Fd_TAI_matSP.rank <- 
  TI.preplot(Fd_PES_matSP.rank)
Fs_TAI_F_matSP.rank <- 
  TI.preplot(Fs_PES_F_matSP.rank) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.rank <- 
  TI.preplot(Fs_PES_M_matSP.rank) %>%
  dplyr::mutate(Sex = "M")

# rlog
Fd_TAI_matSP.rlog <- 
  TI.preplot(Fd_PES_matSP.rlog)
Fs_TAI_F_matSP.rlog <- 
  TI.preplot(Fs_PES_F_matSP.rlog) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.rlog <- 
  TI.preplot(Fs_PES_M_matSP.rlog) %>%
  dplyr::mutate(Sex = "M")
```

```{r}
# maybe I could use geom text? Yes, it is possible but I cannot add the position jitter correctly.
plot_tissue_Fd <- function(preplot_out){
  plot_out <- 
    preplot_out %>%
    ggplot2::ggplot(
      aes(
        x = "Mixed",
        y = TAI,
        group = Tissue
        )
      ) +
    ggplot2::geom_point(
      size = 4, 
      position = ggplot2::position_dodge(width=0.9),
      colour = "#00A087FF") +
    ggplot2::geom_errorbar(
      aes(ymin = TAI - sd, ymax = TAI + sd),
      width = 0.1,
      position = ggplot2::position_dodge(width=0.9),
      colour = "#00A087FF") +
    ggplot2::geom_text(
      aes(label = Tissue),
      vjust = -0.9,
      angle = 90,
      position = ggplot2::position_dodge(width = 0.9)) +
    # ggsci::scale_colour_npg() +
    ggplot2::labs(
      x = "Sex",
      title = "Fucus distichus \n(matSP)"
    ) +
    ggplot2::theme_classic()
  return(plot_out)
}

plot_tissue_Fs <- function(preplot_out){
  plot_out <- 
    preplot_out %>%
    ggplot2::ggplot(
      aes(
        x = Sex,
        y = TAI,
        group = Tissue,
        colour = Sex
        )
      ) +
    ggplot2::geom_point(
      size = 4, 
      position = ggplot2::position_dodge(width=0.9)) +
    ggplot2::geom_errorbar(
      aes(ymin = TAI - sd, ymax = TAI + sd),
      width = 0.1,
      position = ggplot2::position_dodge(width=0.9)) +
    ggplot2::geom_text(
      aes(label = Tissue),
      vjust = -0.9,
      angle = 90,
      position = ggplot2::position_dodge(width = 0.9),
      colour = "black") +
    ggplot2::scale_colour_discrete(
      c("#E64B35FF", "#4DBBD5FF"), 
      guide = "none") +
    ggplot2::labs(
      x = "Sex",
      title = "Fucus serratus \n(matSP)",
      colour = NULL
    ) +
    ggplot2::theme_classic()
  return(plot_out)
}
# https://stackoverflow.com/questions/35654364/ggplot-jitter-geom-errorbar
# `ggplot2::position_jitter(seed = 1)` can also be used.
# https://stackoverflow.com/questions/56816072/position-dodge-and-nudge-y-together
# For colours:
# scales::show_col(ggsci::pal_npg("nrc")(9))
```


```{r, fig.height = 4, fig.width = 2}
plot_tissue_Fd(Fd_TAI_matSP) + 
  ggplot2::labs(subtitle = "TPM")
plot_tissue_Fd(Fd_TAI_matSP.sqrt) + 
  ggplot2::labs(subtitle = "sqrt(TPM)")
plot_tissue_Fd(Fd_TAI_matSP.log2) + 
  ggplot2::labs(subtitle = "log2(TPM+1)")
plot_tissue_Fd(Fd_TAI_matSP.rank) + 
  ggplot2::labs(subtitle = "rank(TPM)")
plot_tissue_Fd(Fd_TAI_matSP.rlog) + 
  ggplot2::labs(subtitle = "rlog(TPM)")
```

```{r, fig.height = 4, fig.width = 3}
base::rbind(Fs_TAI_F_matSP, Fs_TAI_M_matSP) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "TPM")

base::rbind(Fs_TAI_F_matSP.sqrt, Fs_TAI_M_matSP.sqrt) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "sqrt(TPM)")

base::rbind(Fs_TAI_F_matSP.log2, Fs_TAI_M_matSP.log2) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "log2(TPM)")

base::rbind(Fs_TAI_F_matSP.rank, Fs_TAI_M_matSP.rank) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "rank(TPM)")

base::rbind(Fs_TAI_F_matSP.rlog, Fs_TAI_M_matSP.rlog) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "rlog(TPM)")
```

## Perform `FlatLineTest()`

```{r}
Fd_PES_matSP_tS <- 
  tfStability(
    Fd_PES_matSP,
    transforms = c("none", "sqrt", "log2", "rank"),
    permutations = 5000
    )
Fd_PES_matSP_tS_processed <-
  tibble::as_tibble(Fd_PES_matSP_tS, rownames = "transformation")

Fs_PES_F_matSP_tS <- 
  tfStability(
    Fs_PES_F_matSP,
    transforms = c("none", "sqrt", "log2", "rank"),
    permutations = 5000
    )
Fs_PES_F_matSP_tS_processed <-
  tibble::as_tibble(Fs_PES_F_matSP_tS, rownames = "transformation")

Fs_PES_M_matSP_tS <- 
  tfStability(
    Fs_PES_M_matSP,
    transforms = c("none", "sqrt", "log2", "rank"),
    permutations = 5000
    )
Fs_PES_M_matSP_tS_processed <-
  tibble::as_tibble(Fs_PES_M_matSP_tS, rownames = "transformation")

# rlog data

Fd_PES_matSP_tS.rlog <- 
  tfStability(
    Fd_PES_matSP.rlog,
    transforms = c("none"), # it is already transformed.
    permutations = 5000
    )
Fd_PES_matSP_tS.rlog_processed <-
  tibble::as_tibble(Fd_PES_matSP_tS.rlog, rownames = "transformation") %>%
  dplyr::mutate(transformation = "rlog")

Fs_PES_F_matSP_tS.rlog <- 
  tfStability(
    Fs_PES_F_matSP.rlog,
    transforms = c("none"), # it is already transformed.
    permutations = 5000
    )
Fs_PES_F_matSP_tS.rlog_processed <-
  tibble::as_tibble(Fs_PES_F_matSP_tS.rlog, rownames = "transformation") %>%
  dplyr::mutate(transformation = "rlog")

Fs_PES_M_matSP_tS.rlog <- 
  tfStability(
    Fs_PES_M_matSP.rlog,
    transforms = c("none"), # it is already transformed.
    permutations = 5000
    )
Fs_PES_M_matSP_tS.rlog_processed <-
  tibble::as_tibble(Fs_PES_M_matSP_tS.rlog, rownames = "transformation") %>%
  dplyr::mutate(transformation = "rlog")

# combine data

Fd_PES_matSP_tS_res <- 
  dplyr::bind_rows(Fd_PES_matSP_tS_processed, Fd_PES_matSP_tS.rlog_processed) %>%
  dplyr::rename("pval" = value) %>%
  dplyr::mutate(test = "FlatLineTest")

Fs_PES_F_matSP_tS_res <- 
  dplyr::bind_rows(Fs_PES_F_matSP_tS_processed, Fs_PES_F_matSP_tS.rlog_processed) %>%
  dplyr::rename("pval" = value) %>%
  dplyr::mutate(test = "FlatLineTest")

Fs_PES_M_matSP_tS_res <- 
  dplyr::bind_rows(Fs_PES_M_matSP_tS_processed, Fs_PES_M_matSP_tS.rlog_processed) %>%
  dplyr::rename("pval" = value) %>%
  dplyr::mutate(test = "FlatLineTest")

```

_Fucus distichus_

```{r, fig.height = 3.5, fig.width = 5}
Fd_PES_matSP_tS_res %>%
  ggplot2::ggplot(
    aes(
      y = -log10(pval), 
      x = transformation)) +
  ggplot2::geom_col(width = 0.05) +
  ggplot2::geom_point(size = 5) +
  ggplot2::geom_hline(
    yintercept = -log10(0.05), 
    colour = "#E64B35FF",
    linetype = 'dashed',
    size = 2
    ) +
  ggplot2::annotate(
    "label",
    x = 0.7, 
    y = max(max(-log10(Fd_PES_matSP_tS_res$pval))/6, 2.5),
    label = "pval < 0.05",
    # angle = 270,
    colour = "#E64B35FF") +
  ggplot2::labs(
    title = "Fucus distichus matSP",
    subtitle = "FlatLineTest",
    x = "Transformation"
  ) +
  ggplot2::coord_flip() +
  ggplot2::theme_classic()
```

_Fucus serratus_

```{r, fig.height = 3.5, fig.width = 5}
Fs_PES_F_matSP_tS_res %>%
  ggplot2::ggplot(
    aes(
      y = -log10(pval), 
      x = transformation)) +
  ggplot2::geom_col(width = 0.05) +
  ggplot2::geom_point(size = 5) +
  ggplot2::geom_hline(
    yintercept = -log10(0.05), 
    colour = "#E64B35FF",
    linetype = 'dashed',
    size = 2
    ) +
  ggplot2::annotate(
    "label",
    x = 0.7, 
    y = max(max(-log10(Fs_PES_F_matSP_tS_res$pval))/6, 2.5),
    label = "pval < 0.05",
    # angle = 270,
    colour = "#E64B35FF") +
  ggplot2::labs(
    title = "Fucus serratus matSP (female)",
    subtitle = "FlatLineTest",
    x = "Transformation"
  ) +
  ggplot2::coord_flip() +
  ggplot2::theme_classic()

Fs_PES_M_matSP_tS_res %>%
  ggplot2::ggplot(
    aes(
      y = -log10(pval), 
      x = transformation)) +
  ggplot2::geom_col(width = 0.05) +
  ggplot2::geom_point(size = 5) +
  ggplot2::geom_hline(
    yintercept = -log10(0.05), 
    colour = "#E64B35FF",
    linetype = 'dashed',
    size = 2
    ) +
  ggplot2::annotate(
    "label",
    x = 0.7, 
    y = max(max(-log10(Fs_PES_M_matSP_tS_res$pval))/6, 2.5),
    label = "pval < 0.05",
    # angle = 270,
    colour = "#E64B35FF") +
  ggplot2::labs(
    title = "Fucus serratus matSP (male)",
    subtitle = "FlatLineTest",
    x = "Transformation"
  ) +
  ggplot2::coord_flip() +
  ggplot2::theme_classic()
```

# denoised data


Non-transformed

```{r}
# tissues in the matSP
Fd_PES_matSP.denoised <-
  readr::read_csv(file = "data/Fd_PES_matSP_denoised.csv")
Fs_PES_F_matSP.denoised <-
  readr::read_csv(file = "data/Fs_PES_F_matSP_denoised.csv")
Fs_PES_M_matSP.denoised <-
  readr::read_csv(file = "data/Fs_PES_M_matSP_denoised.csv")
```

sqrt-tranformed

```{r}
# tissues in the matSP
Fd_PES_matSP.sqrt.denoised <-
  readr::read_csv(file = "data/Fd_PES_matSP.sqrt_denoised.csv")
Fs_PES_F_matSP.sqrt.denoised <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.sqrt_denoised.csv")
Fs_PES_M_matSP.sqrt.denoised <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.sqrt_denoised.csv")
```

log2-tranformed

```{r}
# tissues in the matSP
Fd_PES_matSP.log2.denoised <-
  readr::read_csv(file = "data/Fd_PES_matSP.log2_denoised.csv")
Fs_PES_F_matSP.log2.denoised <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.log2_denoised.csv")
Fs_PES_M_matSP.log2.denoised <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.log2_denoised.csv")
```

####


```{r}
Fd_TAI_matSP.denoised <- 
  TI.preplot(Fd_PES_matSP.denoised)
Fs_TAI_F_matSP.denoised <- 
  TI.preplot(Fs_PES_F_matSP.denoised) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.denoised <- 
  TI.preplot(Fs_PES_M_matSP.denoised) %>%
  dplyr::mutate(Sex = "M")

# sqrt
Fd_TAI_matSP.sqrt.denoised <- 
  TI.preplot(Fd_PES_matSP.sqrt.denoised)
Fs_TAI_F_matSP.sqrt.denoised <- 
  TI.preplot(Fs_PES_F_matSP.sqrt.denoised) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.sqrt.denoised <- 
  TI.preplot(Fs_PES_M_matSP.sqrt.denoised) %>%
  dplyr::mutate(Sex = "M")

# log2
Fd_TAI_matSP.log2.denoised <- 
  TI.preplot(Fd_PES_matSP.log2.denoised)
Fs_TAI_F_matSP.log2.denoised <- 
  TI.preplot(Fs_PES_F_matSP.log2.denoised) %>%
  dplyr::mutate(Sex = "F")
Fs_TAI_M_matSP.log2.denoised <- 
  TI.preplot(Fs_PES_M_matSP.log2.denoised) %>%
  dplyr::mutate(Sex = "M")
```

```{r}
# maybe I could use geom text? Yes, it is possible but I cannot add the position jitter correctly.
plot_tissue_Fd <- function(preplot_out){
  plot_out <- 
    preplot_out %>%
    ggplot2::ggplot(
      aes(
        x = "Mixed",
        y = TAI,
        group = Tissue
        )
      ) +
    ggplot2::geom_point(
      size = 4, 
      position = ggplot2::position_dodge(width=0.9),
      colour = "#00A087FF") +
    ggplot2::geom_errorbar(
      aes(ymin = TAI - sd, ymax = TAI + sd),
      width = 0.1,
      position = ggplot2::position_dodge(width=0.9),
      colour = "#00A087FF") +
    ggplot2::geom_text(
      aes(label = Tissue),
      vjust = -0.9,
      angle = 90,
      position = ggplot2::position_dodge(width = 0.9)) +
    # ggsci::scale_colour_npg() +
    ggplot2::labs(
      x = "Sex",
      title = "Fucus distichus \n(matSP)"
    ) +
    ggplot2::theme_classic()
  return(plot_out)
}

plot_tissue_Fs <- function(preplot_out){
  plot_out <- 
    preplot_out %>%
    ggplot2::ggplot(
      aes(
        x = Sex,
        y = TAI,
        group = Tissue,
        colour = Sex
        )
      ) +
    ggplot2::geom_point(
      size = 4, 
      position = ggplot2::position_dodge(width=0.9)) +
    ggplot2::geom_errorbar(
      aes(ymin = TAI - sd, ymax = TAI + sd),
      width = 0.1,
      position = ggplot2::position_dodge(width=0.9)) +
    ggplot2::geom_text(
      aes(label = Tissue),
      vjust = -0.9,
      angle = 90,
      position = ggplot2::position_dodge(width = 0.9),
      colour = "black") +
    ggplot2::scale_colour_discrete(
      c("#E64B35FF", "#4DBBD5FF"), 
      guide = "none") +
    ggplot2::labs(
      x = "Sex",
      title = "Fucus serratus \n(matSP)",
      colour = NULL
    ) +
    ggplot2::theme_classic()
  return(plot_out)
}
# https://stackoverflow.com/questions/35654364/ggplot-jitter-geom-errorbar
# `ggplot2::position_jitter(seed = 1)` can also be used.
# https://stackoverflow.com/questions/56816072/position-dodge-and-nudge-y-together
# For colours:
# scales::show_col(ggsci::pal_npg("nrc")(9))
```


```{r, fig.height = 4, fig.width = 2}
plot_tissue_Fd(Fd_TAI_matSP.denoised) + 
  ggplot2::labs(subtitle = "TPM [denoised]")
plot_tissue_Fd(Fd_TAI_matSP.sqrt.denoised) + 
  ggplot2::labs(subtitle = "sqrt(TPM) [denoised]")
plot_tissue_Fd(Fd_TAI_matSP.log2.denoised) + 
  ggplot2::labs(subtitle = "log2(TPM+1) [denoised]")
```

```{r, fig.height = 4, fig.width = 3}
base::rbind(Fs_TAI_F_matSP.denoised, Fs_TAI_M_matSP.denoised) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "TPM [denoised]")

base::rbind(Fs_TAI_F_matSP.sqrt.denoised, Fs_TAI_M_matSP.sqrt.denoised) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "sqrt(TPM) [denoised]")

base::rbind(Fs_TAI_F_matSP.log2.denoised, Fs_TAI_M_matSP.log2.denoised) %>%
  plot_tissue_Fs() + 
  ggplot2::labs(subtitle = "log2(TPM) [denoised]")
```

## Perform `FlatLineTest()`

```{r}
Fd_PES_matSP_tS.denoised <- 
  tfStability(
    Fd_PES_matSP.denoised,
    transforms = c("none", "sqrt", "log2", "rank"),
    permutations = 5000
    )
Fd_PES_matSP_tS_processed.denoised <-
  tibble::as_tibble(Fd_PES_matSP_tS.denoised, rownames = "transformation")

Fs_PES_F_matSP_tS.denoised <- 
  tfStability(
    Fs_PES_F_matSP.denoised,
    transforms = c("none", "sqrt", "log2", "rank"),
    permutations = 5000
    )
Fs_PES_F_matSP_tS_processed.denoised <-
  tibble::as_tibble(Fs_PES_F_matSP_tS.denoised, rownames = "transformation")

Fs_PES_M_matSP_tS.denoised <- 
  tfStability(
    Fs_PES_M_matSP.denoised,
    transforms = c("none", "sqrt", "log2", "rank"),
    permutations = 5000
    )
Fs_PES_M_matSP_tS_processed.denoised <-
  tibble::as_tibble(Fs_PES_M_matSP_tS.denoised, rownames = "transformation")

# combine data

Fd_PES_matSP_tS_res.denoised <- 
  Fd_PES_matSP_tS_processed.denoised %>%
  dplyr::rename("pval" = value) %>%
  dplyr::mutate(test = "FlatLineTest")

Fs_PES_F_matSP_tS_res.denoised <- 
  Fs_PES_F_matSP_tS_processed.denoised %>%
  dplyr::rename("pval" = value) %>%
  dplyr::mutate(test = "FlatLineTest")

Fs_PES_M_matSP_tS_res.denoised <- 
  Fs_PES_M_matSP_tS_processed.denoised %>%
  dplyr::rename("pval" = value) %>%
  dplyr::mutate(test = "FlatLineTest")

```

_Fucus distichus_

```{r, fig.height = 3.5, fig.width = 5}
Fd_PES_matSP_tS_res.denoised %>%
  ggplot2::ggplot(
    aes(
      y = -log10(pval), 
      x = transformation)) +
  ggplot2::geom_col(width = 0.05) +
  ggplot2::geom_point(size = 5) +
  ggplot2::geom_hline(
    yintercept = -log10(0.05), 
    colour = "#E64B35FF",
    linetype = 'dashed',
    size = 2
    ) +
  ggplot2::annotate(
    "label",
    x = 0.7, 
    y = max(max(-log10(Fd_PES_matSP_tS_res$pval))/6, 2.5),
    label = "pval < 0.05",
    # angle = 270,
    colour = "#E64B35FF") +
  ggplot2::labs(
    title = "Fucus distichus matSP [denoised]",
    subtitle = "FlatLineTest",
    x = "Transformation"
  ) +
  ggplot2::coord_flip() +
  ggplot2::theme_classic()
```

_Fucus serratus_

```{r, fig.height = 3.5, fig.width = 5}
Fs_PES_F_matSP_tS_res.denoised %>%
  ggplot2::ggplot(
    aes(
      y = -log10(pval), 
      x = transformation)) +
  ggplot2::geom_col(width = 0.05) +
  ggplot2::geom_point(size = 5) +
  ggplot2::geom_hline(
    yintercept = -log10(0.05), 
    colour = "#E64B35FF",
    linetype = 'dashed',
    size = 2
    ) +
  ggplot2::annotate(
    "label",
    x = 0.7, 
    y = max(max(-log10(Fs_PES_F_matSP_tS_res$pval))/6, 2.5),
    label = "pval < 0.05",
    # angle = 270,
    colour = "#E64B35FF") +
  ggplot2::labs(
    title = "Fucus serratus matSP (female) [denoised]",
    subtitle = "FlatLineTest",
    x = "Transformation"
  ) +
  ggplot2::coord_flip() +
  ggplot2::theme_classic()

Fs_PES_M_matSP_tS_res.denoised %>%
  ggplot2::ggplot(
    aes(
      y = -log10(pval), 
      x = transformation)) +
  ggplot2::geom_col(width = 0.05) +
  ggplot2::geom_point(size = 5) +
  ggplot2::geom_hline(
    yintercept = -log10(0.05), 
    colour = "#E64B35FF",
    linetype = 'dashed',
    size = 2
    ) +
  ggplot2::annotate(
    "label",
    x = 0.7, 
    y = max(max(-log10(Fs_PES_M_matSP_tS_res$pval))/6, 2.5),
    label = "pval < 0.05",
    # angle = 270,
    colour = "#E64B35FF") +
  ggplot2::labs(
    title = "Fucus serratus matSP (male) [denoised]",
    subtitle = "FlatLineTest",
    x = "Transformation"
  ) +
  ggplot2::coord_flip() +
  ggplot2::theme_classic()
```

Get session info.

```{r}
devtools::session_info()
```