---
title: "4. Further evolutionary transcriptomics analyses (addressing reviewers)"
output:
  html_document:
    fig_width: 6
    fig_height: 4
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

Having pre-processed the data, here, I analyse the evolutionary transcriptomics in three brown algal species, across life cycle stages and in different tissues. I use the term `TAI` often here. It stands for "transcriptome age index" (see the R package [`myTAI`](https://drostlab.github.io/myTAI/) for further details). Note: permutations here are set at 500 rather than 50000 to ensure a fast run.

# Reporting the difference in TAI using the `pairwiseTest()`

Here, we examine the TAI across the _Fucus_ embryogenesis. We find a developmental hourglass.

```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(myTAI)
```

### Load PES

non-transformed

```{r warning=FALSE,message=FALSE}
Fd_PES_matSP <-
  readr::read_csv(file = "data/Fd_PES_matSP.csv", show_col_types = FALSE)
Fs_PES_F_matSP <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.csv", show_col_types = FALSE)
Fs_PES_M_matSP <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.csv", show_col_types = FALSE)
```

sqrt-tranformed

```{r warning=FALSE,message=FALSE}
Fd_PES_matSP.sqrt <-
  readr::read_csv(file = "data/Fd_PES_matSP.sqrt.csv", show_col_types = FALSE)
Fs_PES_F_matSP.sqrt <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.sqrt.csv", show_col_types = FALSE)
Fs_PES_M_matSP.sqrt <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.sqrt.csv", show_col_types = FALSE)
```

log2-tranformed

```{r warning=FALSE,message=FALSE}
Fd_PES_matSP.log2 <-
  readr::read_csv(file = "data/Fd_PES_matSP.log2.csv", show_col_types = FALSE)
Fs_PES_F_matSP.log2 <-
  readr::read_csv(file = "data/Fs_PES_F_matSP.log2.csv", show_col_types = FALSE)
Fs_PES_M_matSP.log2 <-
  readr::read_csv(file = "data/Fs_PES_M_matSP.log2.csv", show_col_types = FALSE)
```

## `PairwiseTest()` for matSP reptip vs vegtip

non-tranformed

```{r warning=FALSE,message=FALSE}
Fd_pairwisetest <-
        myTAI::PairwiseTest(
                ExpressionSet = Fd_PES_matSP,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_F_pairwisetest <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_F_matSP,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_M_pairwisetest <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_M_matSP,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
```

sqrt-tranformed

```{r warning=FALSE,message=FALSE}
Fd_pairwisetest.sqrt <-
        myTAI::PairwiseTest(
                ExpressionSet = Fd_PES_matSP.sqrt,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_F_pairwisetest.sqrt <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_F_matSP.sqrt,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_M_pairwisetest.sqrt <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_M_matSP.sqrt,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
```

log2-tranformed

```{r warning=FALSE,message=FALSE}
Fd_pairwisetest.log2 <-
        myTAI::PairwiseTest(
                ExpressionSet = Fd_PES_matSP.log2,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_F_pairwisetest.log2 <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_F_matSP.log2,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_M_pairwisetest.log2 <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_M_matSP.log2,
                modules = list(contrast1 = 2, contrast2 = 4),
                altHypothesis = "greater",
                permutations = 1000)$p.value
```

```{r}
p_res <- data.frame(
        pvalues = c(Fd_pairwisetest, Fs_M_pairwisetest, Fs_F_pairwisetest, 
          Fd_pairwisetest.sqrt, Fs_M_pairwisetest.sqrt, Fs_F_pairwisetest.sqrt, 
          Fd_pairwisetest.log2, Fs_M_pairwisetest.log2, Fs_F_pairwisetest.log2),
        species = c("Fd", "Fs", "Fs", "Fd", "Fs", "Fs", "Fd", "Fs", "Fs"),
        sex = c("mixed", "M", "F", "mixed", "M", "F", "mixed", "M", "F"),
        tf = c("none", "none", "none", 
                           "sqrt", "sqrt", "sqrt", 
                           "log2", "log2", "log2"),
        sample = c("Fd", "Fs_M", "Fs_F", "Fd", "Fs_M", "Fs_F", "Fd", "Fs_M", "Fs_F")
)

p_res |>
        ggplot2::ggplot(aes(x = sample, y = -log10(pvalues), linetype = tf, group = tf)) +
        ggplot2::geom_line() +
        ggplot2::geom_point() +
        ggplot2::geom_hline(yintercept = -log10(0.05), colour = "red", linetype = 2) +
        ggplot2::coord_flip()
```

From all the transformations, it appears that _F. distichus_ reptip is significantly younger than the vegtip (`p = 7.679781e-18` for sqrt-transformation). In _F. serratus_, this is also the case for males (`p = 1.388551e-06` for sqrt-transformation) but not females (`p = 0.0672199` for sqrt-transformation).

## `PairwiseTest()` for matSP reptip vs all

non-tranformed

```{r warning=FALSE,message=FALSE}
Fd_pairwisetest <-
        myTAI::PairwiseTest(
                ExpressionSet = Fd_PES_matSP,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_F_pairwisetest <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_F_matSP,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_M_pairwisetest <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_M_matSP,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
```

sqrt-tranformed

```{r warning=FALSE,message=FALSE}
Fd_pairwisetest.sqrt <-
        myTAI::PairwiseTest(
                ExpressionSet = Fd_PES_matSP.sqrt,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_F_pairwisetest.sqrt <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_F_matSP.sqrt,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_M_pairwisetest.sqrt <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_M_matSP.sqrt,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
```

log2-tranformed

```{r warning=FALSE,message=FALSE}
Fd_pairwisetest.log2 <-
        myTAI::PairwiseTest(
                ExpressionSet = Fd_PES_matSP.log2,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_F_pairwisetest.log2 <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_F_matSP.log2,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
Fs_M_pairwisetest.log2 <-
        myTAI::PairwiseTest(
                ExpressionSet = Fs_PES_M_matSP.log2,
                modules = list(contrast1 = 2, contrast2 = c(1,3,4)),
                altHypothesis = "greater",
                permutations = 1000)$p.value
```

```{r}
p_res <- data.frame(
        pvalues = c(Fd_pairwisetest, Fs_M_pairwisetest, Fs_F_pairwisetest, 
          Fd_pairwisetest.sqrt, Fs_M_pairwisetest.sqrt, Fs_F_pairwisetest.sqrt, 
          Fd_pairwisetest.log2, Fs_M_pairwisetest.log2, Fs_F_pairwisetest.log2),
        species = c("Fd", "Fs", "Fs", "Fd", "Fs", "Fs", "Fd", "Fs", "Fs"),
        sex = c("mixed", "M", "F", "mixed", "M", "F", "mixed", "M", "F"),
        tf = c("none", "none", "none", 
                           "sqrt", "sqrt", "sqrt", 
                           "log2", "log2", "log2"),
        sample = c("Fd", "Fs_M", "Fs_F", "Fd", "Fs_M", "Fs_F", "Fd", "Fs_M", "Fs_F")
)

p_res |>
        ggplot2::ggplot(aes(x = sample, y = -log10(pvalues), linetype = tf, group = tf)) +
        ggplot2::geom_line() +
        ggplot2::geom_point() +
        ggplot2::geom_hline(yintercept = -log10(0.05), colour = "red", linetype = 2) +
        ggplot2::coord_flip()
```

Results differ between transformations. If we go by the **square-root** transformation, which we were reporting in other cases in the paper, it appears that _F. distichus_ reptip is significantly younger than the mean TAI of other tissues (`p = 0.003029956`). In _F. serratus_, this is also the case for males (`p = 0.003790212`) but not females (`p = 0.402316`).

Get session info.

```{r warning=FALSE,message=FALSE}
devtools::session_info()
```